Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE1)

### 2014-04-27 WSSSPE1_report/WSSSPE1_summary_report.tex
small changes in various places - will submit this morning

This report attempts to summarize the various aspects of the workshop.
The remainder of this paper first gives an overview of the process
with which the workshop was put together (\S\ref{sec:process}), then
proceeds with summaries of the two keynote
presentations~(\S\ref{sec:keynotes}), followed by summaries of the
workshop papers grouped by the three thematic workshop panels to which
they were assigned~(\S\ref{sec:devel}-\ref{sec:community}). Three
broader cross-cutting issues surfaced repeatedly, and are discussed
separately~(\S\ref{sec:cross-cutting}), as are use cases for
sustainable software~(\S\ref{sec:use-cases}). The summaries are based
not only on the papers and panel presentations, but also on the many
comments raised in both the onsite and online discussions that
accompanied the workshop elements, as documented by collaborative note
taking during the workshop~\cite{WSSSPE1-google-notes}. We conclude
with issues and lessons learned, as well as plans for future
activities~(\S\ref{sec:conclusions}).  The original call for papers is included in 
Appendix~\ref{sec:cfp}.  The short papers accepted to
the workshop are listed in Appendix~\ref{sec:papers}, and a partial
list of workshop attendees can be found in
Appendix~\ref{sec:attendees}.

@@

Roughly following the call for papers topics, the accepted submissions
were grouped into three main
categories, namely \emph{Developing and Supporting Software},
\emph{Policy}, and \emph{Communities}. Each category was assigned a
panel, with three to four panelists drawn from authors of the
associated submissions, who were each assigned to read and summarily
present a subset of the papers associated with the panel. The process
from organizing and advertising the workshop, to collecting and
reviewing the papers, and putting together the agenda was documented
by the organizers in a report~\cite{WSSSPE1-pre-report}, which they
self-archived in the same way as contributed papers.

@@

In this section, we discuss some of the software projects as case
studies to better understand the points discussed during the workshop
and described in the previous sections. We dive deeper to find how 
they are affected by the sustainability related issues in
practice. Additionally, we discuss issues surrounding people,
especially the career paths of original developers and investigate to
answer the question: \emph{are the original developers still around?}

We classify the software projects discussed in the workshop in two
broad categories. First, the \emph{utility} software, comprising of
general purpose software. Utility software is often used as enabler
or facilitator for the development of other tools and techniques
to carry out scientific work. This includes the software developed to
efficiently utilize new research infrastructures. Second, the
\emph{scientific} software, comprising the software that was
originally developed with an aim to solve a specific scientific
problem.

% Why are we classifying software like above?
This classification is motivated by our argument that the two kinds of
software projects wildly vary in factors such as scope, purpose and
usage. The development and management of each kind is significantly
different. Consequently, the sustainability challenges faced by them
differ and must be treated separately. For instance, the challenges
faced by a gateway software development project such as
\toolname{CIPRES} or visualization software products such as
\toolname{VisIT} or \toolname{ParaView} are distinct to a niche software for
\textit{ab initio} modeling and simulation such as \toolname{VASP} or
\toolname{Quantum Espresso}. %\note{cite SW}

@@

@@

\section*{Developing and Supporting Software}

@@

\section*{Policy}

@@

\section*{Communities, Models, and Education}

>>>>

This report attempts to summarize the various aspects of the workshop.
The remainder of this paper first gives an overview of the process
with which the workshop was organized (\S\ref{sec:process}), then
proceeds with summaries of the two keynote
presentations~(\S\ref{sec:keynotes}), followed by summaries of the
workshop papers grouped by the three thematic workshop panels to which
they were assigned~(\S\ref{sec:devel}-\ref{sec:community}). Three
broader cross-cutting issues surfaced repeatedly, and are discussed
separately~(\S\ref{sec:cross-cutting}), as are use cases for
sustainable software~(\S\ref{sec:use-cases}). The summaries are based
not only on the papers and panel presentations, but also on the many
comments raised in both the onsite and online discussions that
accompanied the workshop elements, as documented by collaborative note
taking during the workshop~\cite{WSSSPE1-google-notes}. We conclude
with issues and lessons learned, as well as plans for future
activities~(\S\ref{sec:conclusions}).  The original call for papers is included in 
Appendix~\ref{sec:cfp}.  The short papers accepted to
the workshop are listed in Appendix~\ref{sec:papers}, and a partial
list of workshop attendees can be found in
Appendix~\ref{sec:attendees}.

@@

Roughly following the call for papers topics, the accepted submissions
were grouped into three main
categories, namely \emph{Developing and Supporting Software},
\emph{Policy}, and \emph{Communities}. Each category was assigned to a
panel, with three to four panelists drawn from authors of the
associated submissions, who were each assigned to read and summarily
present a subset of the papers associated with the panel. The process
from organizing and advertising the workshop, to collecting and
reviewing the papers, and putting together the agenda was documented
by the organizers in a report~\cite{WSSSPE1-pre-report}, which they
self-archived in the same way as contributed papers.

@@

In this section, we discuss some of the software projects as case
studies to better understand the points discussed during the workshop
and described in the previous sections, and to find how 
they are affected by sustainability issues in
practice. 
%Additionally, we discuss issues surrounding people,
%especially the career paths of original developers and investigate to
%answer the question: \emph{are the original developers still around?}

We classify the software projects discussed in the workshop in two
broad categories. First, the \emph{utility} software, comprising of
general purpose software. Utility software is often used as enabler
or facilitator for the development of other tools and techniques
to carry out scientific work. This includes the software developed to
efficiently utilize new research infrastructures. Second, the
\emph{scientific} software, comprising the software that was
originally developed with an aim to solve a specific scientific
problem.
%
% Why are we classifying software like above?
This classification is motivated by our argument that the two kinds of
software projects wildly vary in factors such as scope, purpose and
usage. The development and management of each kind is significantly
different. Consequently, the sustainability challenges faced by them
differ and must be treated separately. For instance, the challenges
faced by a gateway software development project such as
\toolname{CIPRES} or visualization software products such as
\toolname{VisIT} or \toolname{ParaView} are distinct to a niche software for
\textit{ab initio} modeling and simulation such as \toolname{VASP} or
\toolname{Quantum Espresso}. %\note{cite SW}

@@

\subsection{Distinctions}

@@

\subsection*{Developing and Supporting Software}

@@

\subsection*{Policy}

@@

\subsection*{Communities, Models, and Education}
